---
title: "H2 Intro Analysis Summary"
author: "Morgan Browning"
date: "7/20/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: sandstone
    fig_width: 7
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, warning = FALSE, message = FALSE, tidy = TRUE, cache = FALSE)
```
```{r}
## ----Library----------------------------------------

# load all packages into the environment. there are more packages here than I ended up using,
# but the list just kept getting longer over time. could clean this up but doesnt seem worth
# the effort because no harm done
library(relaimpo)
library(plotly)
library(plyr)
library(robustbase)
library(sandwich)
library(lmtest)
library(graphics)
library(colorspace)
library(scales)
library(ggrepel)
library(here)
# library(ggtextures)
library(gridExtra)
library(ggpubr)
library(patternplot)
library(plotly)
library(flexdashboard)
library(DT)
library(naniar)
library(maps)
library(mapdata)
library(usmap)
library(ggmap)
library(maptools)
library(sf)
library(PerformanceAnalytics)
library(jtools)
library(ggstance)
library(huxtable)
library(readxl)
library(knitr)
library(tidyverse)
library(kableExtra)
library(ggthemes)
library(patchwork)
library(here)

## ----Functions-------------------------------------------

# function to pull in all sheets from an excel spreadsheet and turn them into 
# separate data frames. should make for easier data analysis and manipulation 

ReadAllSheets <- function(filename, tibble = FALSE) {
  sheets <- readxl::excel_sheets(filename)
  x <- lapply(sheets, function(X) readxl::read_excel(filename, 
                                                     sheet = X, 
                                                     skip = 6))
  if(!tibble) x <- lapply(x, as.data.frame)
  names(x) <- sheets
  x
}

# function to turn my scenario titles (which are very convoluted and long) into 
# intelligible combinations of variables that I can filter on to better sort the 
# data and results

categorize <- function(table) {
  x <- table %>%
    mutate(h2req = case_when(
      str_detect(Scenario, "5p") ~ "5p",
      str_detect(Scenario, "10p") ~ "10p",
      str_detect(Scenario, "15p") ~ "15p",
      str_detect(Scenario, "20p") ~ "20p",
      str_detect(Scenario, "25p") ~ "25p",
      TRUE ~ "Other"
    )) %>%
    mutate_if(is.numeric, ~round(.,2)) %>%
    mutate_all(~replace(.,is.na(.),0))
}

## ----import data------------------------------------------

# points to the excel file of output from VEDA_BE, reads in the names of the sheets,
# and plugs the spreadsheet into the function I created to turn all sheets into
# separate dataframes in the current environment

results <- c(here("h2_data/h2data_072020.xlsx"))
data_global <- ReadAllSheets(results)

## ----co2 emissions------------------------------------------

co2emissions_reg <- as.data.frame(data_global$`CO2 Emissions`) %>% 
  select(-`2011`,-`Attribute`) %>%
  gather(`2010`, `2015`, `2020`, `2025`, `2030`, `2035`, `2040`, `2045`, `2050`, 
         key = "Year", value = "emissions") %>%
  filter(Commodity != "ETHCO2") %>%
  mutate_if(is.numeric, ~round(.,2)) %>%
  mutate_all(~replace(.,is.na(.),0))

co2emissions <- co2emissions_reg %>%
  group_by(Scenario, Commodity, Year) %>%
  summarize(emissions = sum(emissions))


## ----h2 production------------------------------------------------------

h2prod_cap_reg <- as.data.frame(data_global$`H2 Production Capacity`) %>% 
  gather(`2015`, `2020`, `2025`, `2030`, `2035`, `2040`, `2045`, `2050`, 
         key = "Year", value = "cap") %>% 
  mutate_if(is.numeric, ~round(.,2)) %>%
  mutate_all(~replace(.,is.na(.),0))

h2prod_cumcap_reg <- h2prod_cap_reg %>%
  group_by(Scenario, Year, Attribute, Region) %>%
  summarize(cap = sum(cap)) %>%
  ungroup()

h2prod_cap <- h2prod_cap_reg %>%
  group_by(Scenario, Year, Attribute, Process) %>%
  summarize(cap = sum(cap)) %>%
  ungroup()

h2prod_cumcap <- h2prod_cap %>%
  group_by(Scenario,Year,Attribute) %>%
  summarize(cap = sum(cap)) %>%
  ungroup()

## ----h2 production - fuel use------------------------------------------------------

h2prodfuel_reg <- as.data.frame(data_global$`H2 Production Fuels`) %>% 
  gather(`2015`, `2020`, `2025`, `2030`, `2035`, `2040`, `2045`, `2050`, 
         key = "Year", value = "use") %>% 
  mutate_if(is.numeric, ~round(.,2)) %>%
  mutate_all(~replace(.,is.na(.),0))

h2prodfuel <- h2prodfuel_reg %>%
  group_by(Scenario, Process, Commodity, Year, Attribute) %>%
  summarize(use = sum(use)) %>%
  ungroup()

## ----h2 t&d------------------------------------------------------

h2td_reg <- as.data.frame(data_global$`H2 T&D`) %>%
  gather(`2015`, `2020`, `2025`, `2030`, `2035`, `2040`, `2045`, `2050`, 
         key = "Year", value = "use") %>% 
  mutate_if(is.numeric, ~round(.,2)) %>%
  mutate_all(~replace(.,is.na(.),0))
h2td_reg$Year <- as.numeric(h2td_reg$Year)

h2td <- h2td_reg %>%
  group_by(Scenario, Process, Commodity, Year) %>%
  summarize(use = sum(use)) %>%
  ungroup()

## ----h2 in transportation------------------------------------------------------

h2trn_reg <- as.data.frame(data_global$`H2 Transportation Use`) %>%
  gather(`2015`, `2020`, `2025`, `2030`, `2035`, `2040`, `2045`, `2050`, 
         key = "Year", value = "PJ") %>% 
  mutate_if(is.numeric, ~round(.,2)) %>%
  mutate_all(~replace(.,is.na(.),0))
h2trn_reg$Year <- as.numeric(h2trn_reg$Year)

h2trn <- h2trn_reg %>%
  group_by(Scenario, Commodity, Process, Attribute, Year) %>%
  summarize(PJ = sum(PJ)) %>%
  ungroup()

## ----transportation mix------------------------------------------------------

trn_reg <- as.data.frame(data_global$`Transportation Summary`) %>%
  select(-`2011`) %>%
  gather(`2010`, `2015`, `2020`, `2025`, `2030`, `2035`, `2040`, `2045`, `2050`, 
         key = "Year", value = "PJ") %>% 
  mutate_if(is.numeric, ~round(.,2)) %>%
  mutate_all(~replace(.,is.na(.),0))
trn_reg$Year <- as.numeric(trn_reg$Year)

trn <- trn_reg %>%
  group_by(Scenario, Attribute, ProcessSet, Year) %>%
  summarize(PJ = sum(PJ)) %>%
  ungroup()

## Load Setup and Data Scripts ----

# allows me to quickly run my two previous scripts when I make changes in them or
# if I'm just going to work with results and need to pull in the data

# source("h2_Scripts/h2_setup_072020.R")
# source("h2_Scripts/h2_data_072020.R")

## CO2 Emissions ----

co2emissions_graph <- co2emissions %>% 
  filter(Commodity %in% c("COMCO2", "ELCCO2", "RESCO2", "TRNCO2", "INDCO2")) %>%
  ggplot() +
  geom_line(aes(x = Year, y = emissions, color = Scenario, group = Scenario)) +
  facet_wrap(~Commodity, scales = "free_y") +
  labs(x = "Year", y = "CO2 Emissions (Mt)", title = "CO2 Emissions by Scenario")
  

## H2 Production ----

## ~ Total Capacity ----

# separated by technology

h2prod_totalcap_tech <- h2prod_cap %>% filter(Attribute == "VAR_Cap") %>%
  ggplot() +
  geom_line(aes(x = Year, y = cap, color = Scenario, group = Scenario)) +
  facet_wrap(~Process, scales = "free_y") +
  labs(x = "Year", y = "Total Capacity (PJ)", title = "H2 Total Production by Technology")

# cumulative over all technologies

h2prod_totalcap <- h2prod_cumcap %>% filter(Attribute == "VAR_Cap") %>%
  ggplot() +
  geom_line(aes(x = Year, y = cap, color = Scenario, group = Scenario)) +
  labs(x = "Year", y = "Total Capacity (PJ)", title = "H2 Total Production by Scenario")

## ~ New Capacity ----

# separated by technology

h2prod_newcap_tech <- h2prod_cap %>% filter(Attribute == "VAR_Ncap") %>%
  ggplot() +
  geom_line(aes(x = Year, y = cap, color = Scenario, group = Scenario)) +
  facet_wrap(~Process, scales = "free_y") +
  labs(x = "Year", y = "New Capacity (PJ)", title = "H2 New Production Capacity by Technology")

# cumulative over all technologies

h2prod_newcap <- h2prod_cumcap %>% filter(Attribute == "VAR_Ncap") %>%
  ggplot() +
  geom_line(aes(x = Year, y = cap, color = Scenario, group = Scenario)) +
  labs(x = "Year", y = "New Capacity (PJ)", title = "H2 New Production by Scenario")


## ~ Fuel Use ----

h2fueluse <- h2prodfuel %>% 
  filter(Commodity %in% c("ELCH2OC", "ELCH2OW","INDELC","INDNGA")) %>%
  ggplot() +
  geom_line(aes(x = Year, y = use, color = Scenario, group = Scenario)) +
  facet_grid(Commodity~Process, scales = "free_y") +
  labs(y = "Fuel Use", title = "Fuel Use for H2 Production")

## ~ Production ----

h2prodbytech <- h2prodfuel %>% 
  filter(Commodity == "H2PROD") %>%
  ggplot() +
  geom_line(aes(x = Year, y = use, color = Scenario, group = Scenario)) +
  facet_grid(~Process, scales = "free_y") +
  labs(y = "PJ", title = "H2 Production by Technology")

## T&D----

h2td_graph <- h2td %>%
  ggplot() +
  geom_line(aes(x = Year, y = use, color = Scenario, group = Scenario)) +
  facet_grid(~Process)

## H2 in Transportation----

h2trn_type <- h2trn %>%
  filter(Commodity %in% c("TLH2", "THH2")) %>%
  filter(!Process %in% c("SCTHH2", "SCTLH2")) %>%
  ggplot() +
  geom_line(aes(x = Year, y = PJ, group = Scenario, color = Scenario)) +
  facet_grid(Commodity~Process)
  
## Transportation Mix----

trn_mix <- trn %>%
  filter(Attribute == "VAR_FIn") %>%
  filter(ProcessSet %in% c("TH-DSL", "TH-H2")) %>%
  select(-Attribute) %>%
  ggplot() +
  geom_line(aes(x = Year, y = PJ, color = Scenario, group = Scenario)) +
  facet_wrap(~ProcessSet)
```

## Objective

### Tasks

1. Validate H2 build within TIMES
2. Establish a baseline
3. Test cost sensitiities 
4. Highlight areas of interest for scenario analysis

### Methodology Updates

* T&D
  + Reverted to existing TIMES T&D infrastructure while we continue to investigate updates to T&D costs using HDSAM and other datasets
* Transportation sector H2 options
  + HDV - hauling, busses
  + LDV - compact, full, minivan, crossover truck 
* Scenario prep
  + H2 mandates for transportation sector
  + Emissions Constraints
  + Cost sensitivities along the H2 supply chain
  + Renewable electricity requirements for H2 production
  + Water consumption and withdrawal constraints

## Scenario Summary

*5-25% required H2 fuel use for both LDV and HDV (haul only) demands
  + 1% required by 2030
  + scenario required % by 2050 - linearly increasing from 2030 to 2050

## H2 Production by Technology {.tabset}

H2CCNGA = Natural Gas SMR
H2CFPHOTO4 = Photoelectrochemical

### Total Capacity

```{r}
ggplotly(h2prod_totalcap_tech)
```

### New Capacity

```{r}
ggplotly(h2prod_newcap_tech)
```
 
## H2 Production Capacity {.tabset}

### Total Capacity

```{r}
ggplotly(h2prod_totalcap)
```

### New Capacity

```{r}
ggplotly(h2prod_newcap)
```

## Production Inputs & Outputs {.tabset}

ELCH2OC - consumed H2O
ELCH2OW - withdrawn and returned H2O
INDELC - electricity consumed by the industrial sector
INDNGA - natural gas consumed by the industrial sector

### Fuel Use

```{r}
ggplotly(h2fueluse)
```

### Technology Mix

```{r}
ggplotly(h2prodbytech)
```

## Transmission and Distribution 

PLSVUA and PLTKUA are both pipelines - one for transmission, one for distribution

```{r}
ggplotly(h2td_graph)
```

## Transportation Sector {.tabset}

* Vehicles
  + THH2 - H2 haul
  + TLCH2 - H2 LDV compact 
  + TLFH2 - H2 LDV full
  + TLTCOH2 - H2 LDV crossover truck
  + TLVH2 - H2 LDV minivan

* Fuel Types
  + TH-DSL - diesel used by haulers
  + TH-H2 - H2 used by haulers

### H2 Vehicles 

```{r}
ggplotly(h2trn_type)
```

### HDV Tradeoffs

```{r}
ggplotly(trn_mix)
```









