---
title: "TIMES H2 Calibration and Initial Results"
output: 
  flexdashboard::flex_dashboard:
    theme: lumen
    vertical_layout: scroll
    source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
```

# H2 Production

## By Technology

```{r}
h2prod_tech_all
```

> Natural gas SMR H2 production far surpasses PEM-only production

```{r}
h2prod_tech_nonextgen
h2prod_tech_pem
```

## Total H2

```{r}
h2prod_cum_all
```

> No next gen and grid electricity allows for a cheaper option, natural gas, which elicits more H2 production. PEM-only using grid electricity gets higher H2 production than when renewable electricity is required.

## Regional Production

```{r}
h2prod_tech_reg
h2prod_cum_reg
```

> When PEM-only requirement is in place, all regions produce H2. If natural gas SMR is allowed, region 7 becomes the production hub. This is something that we'll need to address/constrain using inter-state trade parameters and costs.

# Electricity Production

## Total Production

```{r}
totalelc
```

> Transportation sector CO2 caps increase total electricity production. PEM requirement elicits more overall electricity production.

## Regional Production
```{r}
regtotalelc
```

> Transportation sector CO2 caps similarly increase electricity production in all regions, but region 6 is differently affected. We will need to dig into why this is.

## By technology

```{r}
elctechs_all
elctechs
```

> PEM requirement increases electricity production, predominantly natural gas, even though renewable electricity is required as a PEM input. This likely shows a needed total electricity output increase due to the shift in electricity to transportation, not only due to PEM electricity needs.


# CO2 Emissions

## All CO2 Emissions

```{r}
co2emissions_all
co2emissions_PEMonly
co2emissions_nonextgen
```

> Likely that increased electricity need in transportation due to the transportation CO2 caps causes the commercial and industrial sectors to use a greater amount of natural gas in place of electricity, increasing those sectors' emissions. 

## Commercial and Industrial CO2 Emissions

```{r}
comco2emissions_all
indco2emissions_all
```

## Electric Sector CO2 Emissions

```{r}
elcco2emissions_all
```

> Requiring renewable electricity reduces emissions from the electric sector as compared to using grid electricity. Tighter transportation sector constraints increase electric sector emissions, as requirements for electricity increase.

## Transportation CO2 Emissions

```{r}
trnco2emissions_all
```

> As expected, except that we see that disallowing next gen H2 production techs gets us to lower emissions as compared to a PEM requirement in the 80% transportation CO2 cap scenario.

```{r}
trnco2emissions_PEMonly
```

> Reducing the cost of electrolysis from 50% to 80% reductions doesnt decrease overall transportation CO2 emissions.

# SO2 Emissions

## All CO2 Emissions

```{r}
so2emissions_all
so2emissions_change
```

> Transportation 80% CO2 caps increase SO2 emissions in the commercial and industrial sectors. Refinery and resource supply emissions decrease as transportation CO2 emissions caps increase.




# Start

## Column {data-width=500, .tabset}

### Chart A1

```{r}

```

> Monthly deaths from lung disease in the UK, 1974–1979

### Chart A2

```{r}

```

### Chart A3

```{r}

```

## Column {data-width=250}

### ValueBox

```{r}
valueBox(dim(mtcars)[1], icon = "fa-pencil")
```

### Chart C

```{r}
gauge(99, min = 0, max = 100,
      gaugeSectors(success = c(90,100),
                   warning = c(50,89),
                   danger = c(0,49)))
```

## Column {data-width=250}

### Chart D

Hello, this is my report. Btw, I can change colors.
<font color = "red">Hello Red Font</font>

# Expert

```{r}
datatable(mydata,
          extensions = 'Buttons', 
          options = list(
            dom = 'Bfrtip',
            buttons = c('copy', 'print', 'csv')
          ))
```

